What can GitHub and Googlecode Issues tell us about Exceptions in Android?

### 2014-08-31 android-stacks.tex
discussions before wrappings

INSERIR EM ALGUM PONTO DO TEXTO:

@@

Throwing general exceptions, such as direct instances of RuntimeException, is  considered a
bad practice, because the exception type does not carry enough information to identify the
cause of the exceptional behaviour, and as a consequence developers need to relying on
the exception message which can be neither complete nor precise~\cite{gosling2000java}.

@@

To get a broather view of the root causes of stack traces, we performed an empirical evaluation
whose goal was to categorized the top 100 root causes.

After a first inspection of the complete list of exceptions reported, a first set of exception 
categories was defined. This list represent the most common terms reffered on exception
names and javadoc documentation. A similar approach was addopted in [COELHO-ECOOP].
The list of categories is illustrated in Table~\ref{tab:categories}. Hence in the second phase
of the study, each exception was mapped to one of the categories previously defined - 
this analysis was based on the inspection of source code, Javadoc, or additional 
documentation related to the exception.

To ensure the quality of the classification, three independent coders classified a randomly selected
sample of 25 exception types (from the total 100) using the same categories;
the interater agreement was high (96\%). We can therefore assume that the
categories are descriptive enough for covering the whole range of exception
types that can occur in real world Java projects.

\begin{table}
  \centering
  \begin{tabular}{|p{2cm}| p{5cm}|}
    \hline
    \bfseries{Category} & \bfseries{Description} \\
    \hline
      Programming logic &  Exceptions thrown by the JVM when the 
evaluation of an expression violates the normal semantics of language (e.g., 
out-of-bounds array index, division-by-zero, access to a null reference). \\ \hline
      Resources (IO, Memory, Energy)  & Exception releated to the manipulation or restriction of resources (e.g., file, network, memory) \\ \hline
      Security                               & Exception releated to security issues (e.g. password validation, criptography) \\ \hline
      Concurrency                            & Exceptions related to multi-threaded programming \\ \hline
      Backward compatibility                 &  A list of exceptions related to backward compatibility - a complete list can be found here [ref]  \\ \hline
      Reflection                             & Exceptions related to Java reflection library  \\ \hline
      Specific                & Exceptions created for specific applications, frameworks or libraries \\ \hline
      General (Exception, Error, Runtime)    & The top level exceptions in Java    \\ \hline
    \hline
  \end{tabular}
  \caption{Description of categories.}
  \label{tab:categories}
\end{table}

The top 100 root causes correspond to approximately 77\% of all stack traces
reported reported on Java stack-issues and 95\% of all exceptions reported on
Android stack-issues. Table~\ref{tab:tophundrend} illustrates the result of such 
classification.

\begin{table}
  \centering
  \begin{tabular}{lrr}
    \hline
    \bfseries{Category} &  \bfseries{ Ocurrences (\%)} \\
    \hline
      Programming logic (java.lang and util) & 2235 (55,64\%)\\
      Resources (IO)                         &  727 (18,10\%) \\
      Security                               &  165 (4,11\%)\\
      Concurrency                            & 116  (2,89\%)\\
      Backward compatibility                 &  219 (5,45\%) \\
      Reflection                             &  91 (2,27\%)\\
      Specific (GUI,FRAMEWORK)               &  197 (4,90\%)\\
      General (Error, Exception, Runtime)    &  267 (6,65\%)\\
    \hline
  \end{tabular}
  \caption{Characterization of the 100 root causes.}
  \label{tab:tophundrend}
\end{table}

This classification also enable us to find stack traces reporting exceptions 
directly related to concerns often mentioned on research
works on Android Plaftorm, such as: Concurrency, Backward compatibility, Security, etc.
Considering this broader set the programming mistakesstill represent most of the causes (55,64\%)
of the analyzed exception stack traces.

%TODO: Discuss about the interater agreement here...

%\subsection{Sizes of Stack traces}
%Concerning the size of the stacks (i.e. the number of frames composing them), we
%could observe that  sometimes it exceeded 100 frames (158 stack frames).
%Manually inspecting the corresponding execution trace was caused by one of the
%two reasons: recursive calls or exceptions signaled by a method defined deep
%down in a reused framework; exceptions successively wrapped. Irrespective of the
%reason, in exception flows involving many methods, it is practically impossible
%to handle an exception in a way other than exiting the application, as at the
%place the exception occurs there is hardly enough contextual information to
%perform recovery actions.

@@

\subsection{Characterizing the Most Common Root Exceptions Signalers}
<COMPLEMENT.... INCLUDE THE METHODS AND CLASSES RESPONSIBLE FOR SIGNALING
MOST OF THE EXCEPTIONS AND THE EXCEPTIONS SIGNALED BY THEM. SIMILAR TO THE WRAPPINGS TABLES>

\section{RQ2.  What kinds of exception (mis)-use emerge from stack trace analysis? }

To answer this question, besides distilling the information available on stack traces (i.e., exception signalers, root causes, exception wrappings),
we mined the exception types (i.e., RuntimeException, checked Exception or Error) related to each exception reported on stack traces 
using source code and byte code analyses supported by Exception Miner; and finally checked 
this information against the compiled best practices. Doing so, we mined some characteristics 
of exception stack traces that points to best practices violations.

@@

\subsection{What the Types of Root Exceptions can tell us?}
\subsection{Exploring the Information Embeeded on Exception Types}

<falar dos tipos mais comuns.... Runtime e NullPointer???>
<LINKCAR COM ANALISE DOS SUPERTIPOS>
<complementar falando do grafico acima>
\subsubsection{Undocumented runtime exceptions.} After filtering out all the exceptions implicitly 
signaled by  JVM (due to programming mistakes) and inspecting the signaler methods of such exceptions (i.e.,  direct an indirect 
instances of java.lang.RuntimeException). We could observe that only 1 method out of 118 inspected signalers
 (i.e., 0,8\%) documented the explicitly thrown runtime exception in a Javadoc comment; and none of these methods
included the runtime exception  (reported on the stack) as part of the exception interface of the method (i.e., using 
throws clause on method signature). This result is aligned with the results of other  study conducted by from 
Sacramento et al ~\cite{sacramento2006unchecked} which observed that the
runtime exceptions in .Net programs are most often not documented.

@@

<fazer o link com wrappings>

\subsection{What the Exception Wrappings can Tell us?}
\subsection{Exploring the Information Embeeded on Exception Wrappings}
Table~\ref{tab:wrappingandroid} presents the wrappings found in this study for all
stack traces found in Android repositories. As we can see, some of the checked
exceptions were indeed wrapped in runtime exceptions or even errors, while most
of them were not wrapped along the stack trace.

\begin{table*}
  \centering
  \begin{tabular}{llcccccc}
    \hline
    \bfseries{Wrapper Type}  &  \bfseries{Root Cause Type} &  \bfseries{Projects}  &  \bfseries{Occurrences} & \textsf{Android} & \textsf{Java/Libcore} & \textsf{Lib} & \textsf{App}  \\
    \hline
      Runtime &  Checked   & 70 & 114 & 66 &  0  & 19 &  29 \\
      Runtime   &  Error   & 38  & 57 & 50 &  0  & 6 &  1 \\
      Checked &  Runtime   & 12 & 22 & 4 &  0  & 9 &  9 \\
      Checked & Error      & 7 &  8 & 5 &  0  & 0 &  3 \\
      Error & Checked      & 4  &  7 &  &  0  & 7 &  0     \\
      Error & Runtime     & 6  &  13   & 1 &  1  & 0 &  11    \\
    \hline
  \end{tabular}
\caption{Wrappings comprising different exception types.}
\label{tab:wrappingandroid}
\end{table*}

This analysis also revealed unexpected wrappings such as: Error wrapping a Checked 
exception; a Runtime wrapping and Error; and an Error wrapping a Runtime.
Java is the only language that provides a hybrid exception handling mechanism
which offers  different types to represent different exception behaviors (i.e., error, 
runtime and checked) (see Section~\ref{sec:best}). According to Java specification Errors should not be
handled inside the system since they usually represent unrecoverable conditions
detected by the JVM such as OutOfMemoryError. Checked exceptions should
represent recoverable conditions and  Runtime exceptions represent are usually
related to programming mistakes from which the developer is not expected to recover
from.

@@

(1)

@@

 We could find  57 stack traces on 38 projects on which an Error was wrapped in a Runtime Exception.

\begin{table}
\centering
\begin{tabular}{ll}
 \bfseries{Signaler from} & \bfseries{Ocurrences} \\
    \hline
ANDROID	& 50 \\
LIB  & 6 \\
APP	& 1 \\
\hline
  \end{tabular}
\caption{Elements tresponsible for the wrapping.}
\label{tab:wrapping01}
\end{table}

\begin{table}
\centering
\begin{tabular}{ll}
 \bfseries{Signaler method} & \bfseries{Ocurrences} \\
    \hline
android.os.AsyncTask &	25 \\
android.app.ActivityThread & 15 \\
android.view.LayoutInflater & 7 \\
org.asynctasktex.AsyncTaskExd & 3 \\
android.support.v4.content.ModernAsyncTask & 2\\
\hline
  \end{tabular}
\caption{Elements tresponsible for the wrapping.}
\label{tab:wrapping01}
\end{table}

\begin{table}
\centering
\begin{tabular}{lll}
 \bfseries{Method} & \bfseries{Project} & \bfseries{Ocurrences}\\
    \hline
android.os.AsyncTask &	122 & 549  \\
android.view.LayoutInflater & 63 & 163 \\
android.support.v4.* & 100 & 512\\
\hline
  \end{tabular}
\caption{Elements tresponsible for the wrapping.}
\label{tab:wrapping01}
\end{table}

EXAMPLE: android.os.AsyncTask

@@

TO INVESTIGATE: WHICH WERE THE LIBS AND THE APP? AND ITS REPUTATION?

@@

@@

Most of such wrappings however were caused by the reflection Java library, used by Android, Application and library classes.
relaction methods convert every exception thata happens when invoking a reflected method into an INVOCATION TARGET EXCEPTION
this conversion is performed inside native code.

>>>>

@@

Throwing general exceptions, such as direct instances of RuntimeException, violates best
practice III. It is considered a bad practice because exception type does not carry enough
 information to identify the cause of the exceptional behaviour, and as a consequence developers need to relying on
the exception message which can be neither complete nor precise~\cite{gosling2000java}.

@@

To get a broather view of the root causes of stack traces, we performed an empirical evaluation
whose goal was to categorized the top 100 root exceptions. The top 100 root causes correspond 
to approximately 95\% of all exception types reported on the exception stack-issues.

Such categorization was based known sources of faults mobile developement: concurrency ~cite{to_include}
, backward compatibility ~cite{to_include}, security  ~cite{to_include} and resource management (IO, Memory, Batery) ~cite{to_include}.
We added to this set categories to represent the exceptions caused by programming mistakes,
 and general Java exceptions (whose types not convey any specific cause of failure), and 
exceptions that represent specific  exceptional situations of apps, frameworks or libraries.
The list of categories is illustrated in Table~\ref{tab:categories}.

\begin{table}
  \centering
  \begin{tabular}{|p{2cm}| p{5cm}|}
    \hline
    \bfseries{Category} & \bfseries{Description} \\
    \hline
      Programming logic &  Exceptions thrown by the JVM when the 
evaluation of an expression violates the normal semantics of language (e.g., 
out-of-bounds array index, division-by-zero, access to a null reference). \\ \hline
      Resources (IO, Memory, Batery)  & Exception releated to the manipulation or restriction of resources (e.g., file, network, memory) \\ \hline
      Security                               & Exception releated to security issues (e.g. password validation, criptography) \\ \hline
      Concurrency                            & Exceptions related to multi-threaded programming \\ \hline
      Backward compatibility                 &  A list of exceptions related to backward compatibility - a complete list can be found here [ref]  \\ \hline
      Specific Exceptions              & Exceptions created for specific applications, frameworks or libraries \\ \hline
      General (Exception, Error, Runtime)    & The top level exceptions in Java    \\ \hline
    \hline
  \end{tabular}
  \caption{Description of categories.}
  \label{tab:categories}
\end{table}

Each top 100 root exceptions was mapped to one of these categories - 
this analysis was based on the inspection of source code, Javadoc, or additional 
documentation related to each exception.  Table~\ref{tab:tophundrend} 
illustrates the result of such classification, the number of exception stack traces
on which the exceptions of each category was found as the root cause.

\begin{table}
  \centering
  \begin{tabular}{lrr}
    \hline
    \bfseries{Category} &  \bfseries{Occurence on Stacks (\%)} \\
    \hline
      Programming logic (java.lang and util) & 2326 (57,9\%)\\
      Resources (IO)                         &  727 (18,1\%) \\
      Security                               &  165 (4,1\%)\\
      Concurrency                            & 116  (2,9\%)\\
      Backward compatibility                 &  219 (5,5\%) \\
      Specific Exceptions               &  197 (4,9\%)\\
      General (Error, Exception, Runtime)    &  267 (6,7\%)\\
    \hline
  \end{tabular}
  \caption{Characterization of the 100 root causes.}
  \label{tab:tophundrend}
\end{table}

To ensure the quality of the classification, three independent coders classified a randomly selected
sample of 25 exception types (from the total 100) using the same categories;
the interater agreement was high (96\%). We could also observe that
We can therefore assume that the categories are descriptive 
enough for covering the range of exception types analyzed.

In this second analysis the programming mistakes were also the
most mentioned root causes, followed by exceptions related to resource 
management (which was expected since on mobile environment 
resources are usualy scarse). We could observe that there were 
stack traces related to the main concerns of mobile develoment.

%TODO: Discuss about the interater agreement here...
%\subsection{Sizes of Stack traces}
%Concerning the size of the stacks (i.e. the number of frames composing them), we
%could observe that  sometimes it exceeded 100 frames (158 stack frames).
%Manually inspecting the corresponding execution trace was caused by one of the
%two reasons: recursive calls or exceptions signaled by a method defined deep
%down in a reused framework; exceptions successively wrapped. Irrespective of the
%reason, in exception flows involving many methods, it is practically impossible
%to handle an exception in a way other than exiting the application, as at the
%place the exception occurs there is hardly enough contextual information to
%perform recovery actions.

@@

\subsection{What the Types of Root Exceptions can tell us?}
%\subsection{Exploring the Information Embeeded on Exception Types}
%\subsection{The Exception Types and What They can Tell us? }

Besides distilling the information available on stack traces (i.e., exception signalers, root causes, exception wrappings),
we mined the exception types (i.e., RuntimeException, checked Exception or Error) related to each exception reported on stack traces 
using source code and byte code analyses supported by Exception Miner described before.

@@

We could observe that most of the exceptions reported on issues were runtime, 
mainly  signaled on the Android Platform and in the Application code. we could also find 
runtime exceptions signaled from libraries.

In order to evaluate if  best practice IV was being obeyed, we manualy inspected the   
signature of each method signaling a runtime exception, that was defined by third 
party code (lib or Android) and reused by an application.
In this task we filtered out all the exceptions implicitly 
signaled by  JVM (due to programming mistakes) since these exceptions should not be documented on the 
method signature.

We could observe that only 1 method out of 118 inspected signalers
 (i.e., 0,8\%) documented the explicitly thrown runtime exception in a Javadoc comment; and none of these methods
included the runtime exception  (reported on the stack) as part of the exception interface of the method (i.e., using 
throws clause on method signature). This result is aligned with the results of other  study conducted by from 
Sacramento et al ~\cite{sacramento2006unchecked} which observed that the
runtime exceptions in .Net programs are most often not documented.

@@

We could also observe checked exceptions (which according to Java specification represent
recoverable conditions) as the root causes of reported exception stack traces.
A checked exception can remain uncaught if it is declared by evey method on the call chain or if 
it is wrapped in a runtime exception and rethrown. Next, section describes the kinds of 
wrappings found in this study.

\subsection{What the Cross-Type  Exception Wrappings can Tell us?}

\subsection{Exploring the Information Embeeded on Exception Wrappings}
Table~\ref{tab:wrappingandroid} presents the wrappings found in this study for all
stack traces found in Android repositories. As we can see, some of the checked
exceptions were indeed wrapped in runtime exceptions or even errors.

%, while most of them were not wrapped along the stack trace.

\begin{table*}
  \centering
  \begin{tabular}{llcccccc}
    \hline
    \bfseries{Wrapper Type}  &  \bfseries{Root Cause Type} &  \bfseries{Projects}  &  \bfseries{Occurrences} & \textsf{Android} & \textsf{Java/Libcore} & \textsf{Lib} & \textsf{App}  \\
    \hline
      Runtime &  Checked   & 70 & 114 & 66 &  0  & 19 &  29 \\
      Runtime   &  Error   & 38  & 57 & 50 &  0  & 6 &  1 \\
      Checked &  Runtime   & 12 & 22 & 4 &  0  & 9 &  9 \\
      Checked & Error      & 7 &  8 & 5 &  0  & 0 &  3 \\
      Error & Checked      & 4  &  7 &  &  0  & 7 &  0     \\
      Error & Runtime     & 6  &  13   & 1 &  1  & 0 &  11    \\
    \hline
      %none   & Runtime   &  2360 & 359 \\
      %none  &   Checked  & 422  & 117 \\
      %none  &   Error   & 381  & 141 \\
      %none &  Undefined  &  15   & 10  \\
      %none & Throwable  &  1    & 1 \\
     %\hline
      %Runtime   & Runtime & 560  & 182 \\    
      %Checked   & Checked & 98   & 42  \\
      %Error     & Error   & 15   & 14  \\
    %\hline
  \end{tabular}
\caption{Wrappings comprising different exception types.}
\label{tab:wrappingandroid}
\end{table*}

This analysis revealed unexpected wrappings such as: Error wrapping a Checked 
exception; a Runtime wrapping and Error; and an Error wrapping a Runtime.
Java is the only language that provides a hybrid exception handling mechanism
which offers  different types to represent different exception behaviors (i.e., error, 
runtime and checked) (see Section~\ref{sec:best}). According to Java specification Errors should not be
handled inside the system since they usually represent unrecoverable conditions
detected by the JVM such as OutOfMemoryError. Checked exceptions should
represent recoverable conditions and  Runtime exceptions represent are usually
related to programming mistakes from which the developer is not expected to recover
from. Next sections discusses about such types of wrappings.

@@

@@

We could find  57 stack traces on 38 projects on which an Error was wrapped in a runtime exception.
Manually inspecting each exception stack trace, we could observe that 44% of the wrappings were performed 
by android.os.AsyncTask.

@@

%TO INVESTIGATE: WHICH WERE THE LIBS AND THE APP? AND ITS REPUTATION?

@@

COMPLETE

@@

Most of such wrappings however were caused by the reflection Java library, used by Android, 
Application and library classes. Reflaction methods convert every exception thata happens 
when invoking a reflected method into an InvoncationTargetException (this conversion is performed inside native code).
